Ethnicity sensitive author disambiguation using semi-supervised learning

### 2016-04-26 paper/author_disambiguation.tex
kesw: first version. 16 pages
Signed-off-by: Mateusz Susik <mateuszsusik@gmail.com>

\usepackage{xspace}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{cite}
\usepackage[hidelinks]{hyperref}
\newcommand*{\eg}{e.g.\@\xspace}
\newcommand*{\ie}{i.e.\@\xspace}
\usepackage{color}
\setlength{\intextsep}{10pt plus 2pt minus 2pt}
\newcommand{\msnote}[1]{\textcolor{blue}{[MS: #1]}}

\def\baselinestretch{0.95}
\begin{document}

@@

@@

Training data is usually not easily available, therefore unsupervised approaches propose
the use of domain-specific, manually designed, linkage functions tailored towards author
disambiguation \cite{malin2005unsupervised,song2007efficient, kang2009co,schulz2014exploiting}.
These approaches have the advantage of not requiring hand-labeled data, but generally do
not perform as well as supervised approaches.
To reconcile both worlds, semi-supervised methods make use of small, manually verified clusters of
publications and/or high-precision domain-specific rules to build a training
set of pairs of publications, from which a linkage function is then built using supervised learning
\cite{ferreira2010effective,torvik2009author,levin2012citation}.

Semi-supervised approaches also allow for the tuning of the
clustering algorithm when the latter is applied to a mixed set of labeled
and unlabeled publications, \eg, by maximizing some clustering performance
metric on the known clusters \cite{levin2012citation}.

@@

In this context, we position the work in this paper as a
semi-supervised solution for author disambiguation, with the significant
advantage of having a very large collection of more than 1 million crowdsourced annotations
of publications whose true authors are identified.
The extent and coverage of this data allows us to revisit, validate and nuance previous
findings regarding supervised learning of linkage functions, and to better explore strategies
for semi-supervised clustering.
Furthermore, by releasing our data in the public domain, we hope to provide a benchmark on
which further research on author disambiguation and related topics can be built.

@@

@@

\subsubsection{Input features.} Following previous works, pairs of signatures $(s_1, s_2)$ are first transformed to vectors $v \in \mathbb{R}^p$
by building so-called similarity profiles \cite{treeratpituk2009disambiguating} on which supervised learning is carried out.
In this work, we design and evaluate fifteen standard input
features based on the comparison of signature fields, as reported in the first
half of Table~\ref{table:features}.
As an illustrative example, the \textit{Full name} feature corresponds to the similarity between the (full)
author name fields of the two signatures, as measured using as combination
operator the cosine similarity between their respective $(n,m)$-\emph{TF-IDF} vector
representations\footnote{$(n,m)$ denotes that the \emph{TF-IDF} vectors are computed
from character $n$, $n+1$, ..., $m$-grams.
When not specified, \emph{TF-IDF} vectors are otherwise computed from words.}.

Author names from different cultures, origins or ethnic groups are likely to be
disambiguated using different strategies (\eg, pairs of signatures with French
author names versus pairs of signatures with Chinese author names) \cite{treeratpituk2012name, chin2014effective}.
To support our disambiguation algorithm, we added seven features to our feature set, with each
evaluating the degree of belonging of both signatures to an ethnic group. \msnote{Explain here ethnicity features (ethnicities come from some kind of US Census. Name those groups (as I removed them from the table))}

More specifically, using census data extracted from \cite{rugglesintegrated},
we build a support vector machine classifier (using a linear kernel and
one-versus-all classification scheme) for mapping the
 $(1,5)$-TF-IDF representation of an author name to one of the seven ethnic groups. Given a
pair of signatures $(s_1, s_2)$, the proposed ethnicity features are each
computed as the estimated probability of $s_1$ belonging to the corresponding
ethnic group, multiplied by the estimated probability of $s_2$ belonging to the
same group. In doing so, the expectation is for the linkage function to become
sensitive to the actual origin of the authors depending on the values of these
features. Indirectly, let us also note that these features hold discriminative
power since if author names are predicted to belong to different ethnic groups,
then they are also likely to correspond to distinct people.

\begin{table*}
\caption{Input features for learning a linkage function}
\label{table:features}
\centering
{\small
\begin{tabular}{|l|l|}
  \hline
  \textbf{Feature} & \textbf{Combination operator}\\
  \hline
  \hline
  Full name & Cosine similarity of $(2,4)$-TF-IDF\\
  Given names & Cosine similarity of $(2,4)$-TF-IDF\\
  First given name & Jaro-Winkler distance\\
  Second given name & Jaro-Winkler distance\\
  Given name initial & Equality\\
  Affiliation & Cosine similarity of $(2,4)$-TF-IDF\\
  Co-authors & Cosine similarity of TF-IDF\\
  Title & Cosine similarity of $(2,4)$-TF-IDF\\
  Journal & Cosine similarity of $(2,4)$-TF-IDF\\
  Abstract & Cosine similarity of TF-IDF\\
  Keywords & Cosine similarity of TF-IDF\\
  Collaborations & Cosine similarity of TF-IDF\\
  References & Cosine similarity of TF-IDF\\
  Subject & Cosine similarity of TF-IDF\\
  Year difference & Absolute difference\\
  \hline
  Any ethnicity feature & Product of estimated probabilities\\
  \hline
\end{tabular}}
\end{table*}

@@

\textit{Building a training set.} The distinctive aspect of our work is the
knowledge of more than 1 million crowdsourced annotations ${\cal C}^\prime = \{
c_0^\prime, ..., c_{M-1}^\prime \}$, indicating together that all signature $s \in
c_i^\prime$ are known to correspond to the same individual $a_i$.
In particular, this data can be used to generate positive pairs $(x=(s_1, s_2), y=0)$ for all
$s_1, s_2 \in c_i^\prime$, for all $i$. Similarly, negative pairs $(x=(s_1,
s_2), y=1)$ can be extracted for all $s_1 \in c_i^\prime, s_2 \in c_j^\prime$, for
all $i \neq j$.

@@

The last component of our author disambiguation pipeline is clustering, that is
the process of grouping together, within a block, all signatures from the same
individual (and only those).
As for many other works on author disambiguation, we make use of hierarchical clustering \cite{ward1963hierarchical} for
building clusters of signatures in a bottom-up fashion.
The method involves iteratively merging together the two most similar clusters until all clusters
are merged together at the top of the hierarchy.
Similarity between clusters is evaluated using either complete, single or average linkage, using
as a pseudo-distance metric the probability that $s_1$ and $s_2$ correspond to
distinct authors, as calculated from the custom linkage function $d$ from Section \ref{methods:linkage}.

@@

As part of this work, we developed a stand-alone application for author
disambiguation, publicly available online\footnote{\url{https://github.com/glouppe/beard}} for free reuse
or study.
Our implementation builds upon the Python scientific stack, making
use of the Scikit-Learn library \cite{scikitlearn} for the supervised learning
of a linkage function and of SciPy for clustering.
All components of the disambiguation pipeline have been designed to follow the
Scikit-Learn API \cite{scikitlearnAPI}, making them easy to maintain,
understand and reuse.
Our implementation is made to be efficient, exploiting parallelization when available, and ready for production environments.
It is also designed to be runnable in an incremental fashion in which our approach is considered to be scalable. We adopt a divide-and-conquer strategy for clustering (called “blocking” in this line of work) in order to reduce the computational complexity from $O(N^2)$ to $O(\sum N_i^2)$, which in practice tends to ${O(N)}$ when $N_i << N$. This also means that instead of having to run the disambiguation process on the whole signature set, the process could be run only on specified blocks if desired.

@@

All the solutions proposed in this work sare evaluated on data extracted from
the \emph{INSPIRE} portal \cite{gentil2009information}, a digital library for scientific literature in
high-energy physics.
Overall, the portal holds more than 1 million publications ${\cal P}$,
forming in total a set ${\cal S}$ of more than 10 million signatures.
Out of these, around 13\% have been \textit{claimed} by their
original authors, marked as such by professional curators or automatically assigned to their true authors thanks
to persistent identifiers provided by publishers or other sources.
Together, they constitute a trusted set $({\cal S}^\prime, {\cal C}^\prime)$ of 15388 distinct individuals sharing
36340 unique author names spread within 1201763 signatures on 360066
publications. This data covers several decades in time and dozens of author
nationalities worldwide.

@@

\begin{table*}
\caption{Average precision, recall and f-measure scores on test folds. Underlined components correspond to the state-of-the-art choices.}
\label{table:results}
\centering
\begin{tabular}{|l|c c c | c c c|}
  \hline
                       & \multicolumn{3}{|c|}{\textbf{B3}} & \multicolumn{3}{|c|}{\textbf{Pairwise}}\\
  \textbf{Description} & $P$ & $R$ & $F$ & $P$ & $R$ & $F$ \\
  \hline
  \hline
Baseline & 0.9024 & 0.9828 & 0.9409  & 0.8298 & 0.9776 & 0.8977 \\
\hline
\underline{Blocking = SFI} & 0.9901 & 0.9760 & 0.9830  & 0.9948 & 0.9738 & 0.9842\\
Blocking = Double metaphone & 0.9856 & 0.9827 & 0.9841  & 0.9927 & 0.9817 & 0.9871 \\
Blocking = NYSIIS & 0.9875 & 0.9826 & \textbf{0.9850}  & 0.9936 & 0.9814 & \textbf{0.9875} \\
Blocking = Soundex & 0.9886 & 0.9745 & 0.9815 & 0.9935 & 0.9725 & 0.9828 \\
\hline
No name normalization & 0.9887 & 0.9697 & 0.9791 & 0.9931 & 0.9658 & 0.9793 \\
\underline{Name normalization} & 0.9901 & 0.9760 & \textbf{0.9830} & 0.9948 & 0.9738 & \textbf{0.9842}  \\
\hline
\hline
\underline{Classifier = GBRT} & 0.9901 & 0.9760 & 0.9830   & 0.9948 & 0.9738 & 0.9842 \\
Classifier = Random Forests & 0.9909 & 0.9783 & \textbf{0.9846}  & 0.9957 & 0.9752 & \textbf{0.9854} \\
Classifier = Linear Regression & 0.9749 & 0.9584 & 0.9666 & 0.9717 & 0.9569 & 0.9643 \\
\hline
Training pairs = Non-blocked, uniform & 0.9793 & 0.9630 & 0.9711  & 0.9756 & 0.9629 & 0.9692 \\
Training pairs = Blocked, uniform & 0.9854 & 0.9720 & 0.9786  & 0.9850 & 0.9707 & 0.9778 \\
\underline{Training pairs = Blocked, balanced} & 0.9901 & 0.9760 & \textbf{0.9830}  & 0.9948 & 0.9738 & \textbf{0.9842} \\
\hline
\hline
\underline{Clustering = Average linkage} & 0.9901 & 0.9760 & \textbf{0.9830}  & 0.9948 & 0.9738 & \textbf{0.9842} \\
Clustering = Single linkage & 0.9741 & 0.9603 & 0.9671  & 0.9543 & 0.9626 & 0.9584 \\
Clustering = Complete linkage & 0.9862 & 0.9709 & 0.9785  & 0.9920 & 0.9688 & 0.9803 \\
\hline
No cut (baseline) & 0.9024 & 0.9828 & 0.9409  & 0.8298 & 0.9776 & 0.8977 \\
Global cut & 0.9892 & 0.9737 & 0.9814  & 0.9940 & 0.9727 & 0.9832 \\
\underline{Block cut} & 0.9901 & 0.9760 & \textbf{0.9830}  & 0.9948 & 0.9738 & \textbf{0.9842} \\
\hline
\hline
Combined best settings & 0.9897 & 0.9827 & \textbf{0.9862}  & 0.9954 & 0.9805 & \textbf{0.9879} \\
  \hline
\end{tabular}
\end{table*}

\textit{Baseline.} The simplest baseline against which we compare our results
consists in grouping all signatures sharing the same (normalized) surname(s) and
the same (normalized) first given name initial. It provides a simple and fast
solution yielding decent results, as reported at the top of
Table~\ref{table:results}

\textit{State-of-the-art.} Most methods proposed in related works have released
neither their software, nor their data, making a fair comparison very difficult.
Yet, we believe solutions reported in the literature can be closely
matched to our generic pipeline, provided the blocking strategy, the linkage
function and the clustering algorithm are properly aligned. In particular,
we consider hereon as the \textit{state-of-the-art} solution the following
combination of components:

@@

As Table~\ref{table:results} shows, switching to either Double metaphone or NYSIIS phonetic-based blocking allows
to improve the overall F-measure score, trading precision for recall.
In particular, the NYSIIS-based phonetic blocking shows to be the most effective when applied
to the state-of-the-art (with an F-measure of $0.9850$) while also being the most
efficient computationally (with 10857 blocks versus 12978 for the baseline).

@@

@@

Clustering together all signatures from the same block (\ie, baseline) is the least effective
strategy ($0.9409$), but yields anyhow surprisingly decent accuracy, given the
fact it requires almost no computation (\ie, both learning a linkage function
and running agglomerative clustering can be skipped -- only the blocking
function is needed to group signatures). In particular, this result reveals
that author names are not ambiguous in most cases\footnote{This holds for the
data we extracted, but may in the future, with the rise of non-Western
researchers, be an underestimate of the ambiguous cases.} and that only a small
fraction of them requires advanced disambiguation procedures. On the other
hand, both global and block cut thresholding strategies give very good results,
with a slight advantage for block cuts ($0.9814$ versus $0.9830$), as expected.
In case ${\cal S}^\prime_b$ is empty (\eg, because it corresponds to a young
researcher at the beginning of his career), this therefore suggests that either using
a cut-off threshold learned globally from the known data or using SFI would in general give
satisfactory results, only marginally worse than if claimed signatures had been
known. Trying out 3 clustering methods, the results show that \textit {average linkage} was the best one.

@@

>>>>

\usepackage{xspace}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{cite}
\usepackage[hidelinks]{hyperref}
\newcommand*{\eg}{e.g.\@\xspace}
\newcommand*{\ie}{i.e.\@\xspace}
\usepackage{color}
\usepackage{microtype}
\setlength{\intextsep}{5pt plus 0pt minus 0pt}
\newcommand{\msnote}[1]{\textcolor{blue}{[MS: #1]}}

\def\baselinestretch{0.98}

\newcommand{\longpage}{\enlargethispage{\baselineskip}}

\begin{document}

@@

\longpage

@@

Training data is usually not easily available, therefore unsupervised approaches propose
the use of domain-specific, manually designed, linkage functions tailored towards author
disambiguation \cite{malin2005unsupervised,song2007efficient, kang2009co,schulz2014exploiting}.
These approaches have the advantage of not requiring hand-labeled data, but generally do
not perform as well as supervised approaches.
To reconcile both worlds, semi-supervised methods make use of small, manually verified clusters of
publications and/or high-precision domain-specific rules to build a training
set of pairs of publications, from which a linkage function is then built using supervised learning
\cite{ferreira2010effective,torvik2009author,levin2012citation}. Semi-supervised approaches also allow for the tuning of the
clustering algorithm when the latter is applied to a mixed set of labeled
and unlabeled publications, \eg, by maximizing some clustering performance
metric on the known clusters \cite{levin2012citation}.

@@

In this context, we position this work as a
semi-supervised solution for author disambiguation, with the significant
advantage of having a very large collection of more than 1 million crowdsourced annotations
of publications whose true authors are identified.
The extent and coverage of this data allows us to revisit, validate and nuance previous
findings regarding supervised learning of linkage functions, and to better explore strategies
for semi-supervised clustering.
Furthermore, by releasing our data in the public domain, we provide a benchmark on
which further research on author disambiguation and related topics can be evaluated.

@@

The distinctive aspect of our work is the
knowledge of more than 1 million crowdsourced annotations, indicating together that all signature $s \in
c_i^\prime$ are known to correspond to the same individual $a_i$.

\longpage

@@

\subsubsection{Input features.} Following previous works, pairs of signatures $(s_1, s_2)$ are first transformed to vectors $v \in \mathbb{R}^p$
by building so-called similarity profiles \cite{treeratpituk2009disambiguating} on which supervised learning is carried out.
In this work, we design and evaluate fifteen standard input
features based on the comparison of signature fields, as reported in the first
half of Table~\ref{table:features}.
As an illustrative example, the \textit{Full name} feature corresponds to the similarity between the (full)
author name fields of the two signatures, as measured using as combination
operator the cosine similarity between their respective $(n,m)$-\emph{TF-IDF} vector
representations\footnote{$(n,m)-$\emph{TF-IDF} vectors are \emph{TF-IDF} vectors computed
from $n$, $n+1$, ..., $m$-grams.}.

Authors from different origins or ethnic groups are likely to be
disambiguated using different strategies (\eg, pairs of signatures with French
author names versus pairs of signatures with Chinese author names)
\cite{treeratpituk2012name, chin2014effective}. For example, scientist coming from China
might more/less often change affiliations, and this dependency, if learnt by the classifier, 
should improve the fit.
To support our disambiguation algorithm, we added seven features to our feature set, with each
evaluating the degree of belonging of both signatures to an ethnic group.

\longpage

More specifically, using census data extracted from \cite{rugglesintegrated},
we build a support vector machine classifier (using a linear kernel and
one-versus-all classification scheme) for mapping the
$(1,5)$-TF-IDF representation of an author name to one of the ethnic groups, as
defined in United States federal censuses. These groups are: \textit{White},
\textit{Black of African American}, \textit{American Indian and Alaska Native},
\textit{Asian}, \textit{Native Hawaiian and Other Pacific Islander},
\textit{Japanese}, \textit{Chinese}, \textit{Others}. Given a
pair of signatures $(s_1, s_2)$, the proposed ethnicity features are each
computed as the estimated probability of $s_1$ belonging to the corresponding
ethnic group, multiplied by the estimated probability of $s_2$ belonging to the
same group. Each of the seven races is used to create a single new feature.
In doing so, the expectation is for the linkage function to become
sensitive to the actual origin of the authors depending on the values of these
features. Indirectly, these features also hold discriminative
power since if author names are predicted to belong to different ethnic groups,
then they are also likely to correspond to distinct people.

\begin{table*}
\caption{Input features for learning a linkage function}
\label{table:features}
\centering
{\small
\begin{tabular}{|l|l|}
  \hline
  \textbf{Feature} & \textbf{Combination operator}\\
  \hline
  \hline
  Full name & Cosine similarity of $(2,4)$-TF-IDF\\
  Given names & Cosine similarity of $(2,4)$-TF-IDF\\
  First given name & Jaro-Winkler distance\\
  Second given name & Jaro-Winkler distance\\
  Given name initial & Equality\\
  Affiliation & Cosine similarity of $(2,4)$-TF-IDF\\
  Co-authors & Cosine similarity of TF-IDF\\
  Title & Cosine similarity of $(2,4)$-TF-IDF\\
  Journal & Cosine similarity of $(2,4)$-TF-IDF\\
  Abstract & Cosine similarity of TF-IDF\\
  Keywords & Cosine similarity of TF-IDF\\
  Collaborations & Cosine similarity of TF-IDF\\
  References & Cosine similarity of TF-IDF\\
  Subject & Cosine similarity of TF-IDF\\
  Year difference & Absolute difference\\
  \hline
  Any ethnicity feature & Product of probabilities estimated by SVM\\
  \hline
\end{tabular}}
\end{table*}

@@

\textit{Building a training set.} 1 million of crowdsourced annotations (see Section \ref{methods}) can be used to generate positive pairs $(x=(s_1, s_2), y=0)$ for all
$s_1, s_2 \in c_i^\prime$, for all $i$. Similarly, negative pairs $(x=(s_1,
s_2), y=1)$ can be extracted for all $s_1 \in c_i^\prime, s_2 \in c_j^\prime$, for
all $i \neq j$.

@@

\longpage

The last component of our author disambiguation pipeline is clustering - 
the process of grouping together, within a block, all signatures from the same
individual (and only those).
As for many other works on author disambiguation, we make use of hierarchical clustering \cite{ward1963hierarchical} for
building clusters of signatures in a bottom-up fashion.
The method involves iteratively merging together the two most similar clusters until all clusters
are merged together at the top of the hierarchy.
Similarity between clusters is evaluated using either complete, single or average linkage, using
as a pseudo-distance metric the probability that $s_1$ and $s_2$ correspond to
distinct authors, as calculated from the custom linkage function $d$ from Section \ref{methods:linkage}.

@@

\longpage

As part of this work, we developed a stand-alone application for author
disambiguation, publicly available online\footnote{\url{https://github.com/glouppe/beard}} for free reuse
or study.
Our implementation builds upon the Python scientific stack, making
use of the Scikit-Learn library \cite{scikitlearn} for the supervised learning
of a linkage function and of SciPy for clustering.
All components of the disambiguation pipeline have been designed to follow the
Scikit-Learn API \cite{scikitlearnAPI}, making them easy to maintain,
understand and reuse.
Our implementation is made to be efficient, exploiting parallelization when available, and ready for production environments.
It is also designed to be runnable in an incremental fashion in which our approach is considered to be scalable. We adopt the blocking phase in order to reduce the computational complexity from $O(N^2)$ to $O(\sum N_i^2)$, which in practice tends to ${O(N)}$ when $N_i << N$. This also means that instead of having to run the disambiguation process on the whole signature set, the process could be run only on specified blocks if desired.

@@

All the solutions proposed in this work are evaluated on data extracted from
the \emph{INSPIRE} portal \cite{gentil2009information}, a digital library for scientific literature in
high-energy physics.
Overall, the portal holds more than 1 million publications ${\cal P}$,
forming in total a set ${\cal S}$ of more than 10 million signatures.
Out of these, around 13\% have been \textit{claimed} by their
original authors, marked as such by professional curators or automatically assigned to their true authors thanks
to persistent identifiers provided by publishers or other sources.
Together, they constitute a trusted set $({\cal S}^\prime, {\cal C}^\prime)$ of 15388 distinct individuals sharing
36340 unique author names spread within 1201763 signatures on 360066
publications. This data covers several decades in time and dozens of author
nationalities worldwide.

@@

\longpage

\subsubsection{Baseline.} The simplest baseline against which we compare our results
consists in grouping all signatures sharing the same (normalized) surname(s) and
the same (normalized) first given name initial. It provides a simple and fast
solution yielding decent results, as reported at the top of
Table~\ref{table:results}

\begin{small}
\begin{table*}
\caption{Average precision, recall and f-measure scores on test folds. Underlined components correspond to the state-of-the-art choices.}
\label{table:results}
\centering
\begin{tabular}{|l|c c c | c c c|}
  \hline
                       & \multicolumn{3}{|c|}{\textbf{B3}} & \multicolumn{3}{|c|}{\textbf{Pairwise}}\\
  \textbf{Description} & $P$ & $R$ & $F$ & $P$ & $R$ & $F$ \\
  \hline
  \hline
Baseline & 0.9024 & 0.9828 & 0.9409  & 0.8298 & 0.9776 & 0.8977 \\
\hline
\underline{Blocking = SFI} & 0.9901 & 0.9760 & 0.9830  & 0.9948 & 0.9738 & 0.9842\\
Blocking = Double metaphone & 0.9856 & 0.9827 & 0.9841  & 0.9927 & 0.9817 & 0.9871 \\
Blocking = NYSIIS & 0.9875 & 0.9826 & \textbf{0.9850}  & 0.9936 & 0.9814 & \textbf{0.9875} \\
Blocking = Soundex & 0.9886 & 0.9745 & 0.9815 & 0.9935 & 0.9725 & 0.9828 \\
\hline
No name normalization & 0.9887 & 0.9697 & 0.9791 & 0.9931 & 0.9658 & 0.9793 \\
\underline{Name normalization} & 0.9901 & 0.9760 & \textbf{0.9830} & 0.9948 & 0.9738 & \textbf{0.9842}  \\
\hline
\underline{Classifier = GBRT} & 0.9901 & 0.9760 & 0.9830   & 0.9948 & 0.9738 & 0.9842 \\
Classifier = Random Forests & 0.9909 & 0.9783 & \textbf{0.9846}  & 0.9957 & 0.9752 & \textbf{0.9854} \\
Classifier = Linear Regression & 0.9749 & 0.9584 & 0.9666 & 0.9717 & 0.9569 & 0.9643 \\
\hline
Training pairs = Non-blocked, uniform & 0.9793 & 0.9630 & 0.9711  & 0.9756 & 0.9629 & 0.9692 \\
Training pairs = Blocked, uniform & 0.9854 & 0.9720 & 0.9786  & 0.9850 & 0.9707 & 0.9778 \\
\underline{Training pairs = Blocked, balanced} & 0.9901 & 0.9760 & \textbf{0.9830}  & 0.9948 & 0.9738 & \textbf{0.9842} \\
\hline
\underline{Clustering = Average linkage} & 0.9901 & 0.9760 & \textbf{0.9830}  & 0.9948 & 0.9738 & \textbf{0.9842} \\
Clustering = Single linkage & 0.9741 & 0.9603 & 0.9671  & 0.9543 & 0.9626 & 0.9584 \\
Clustering = Complete linkage & 0.9862 & 0.9709 & 0.9785  & 0.9920 & 0.9688 & 0.9803 \\
\hline
No cut (baseline) & 0.9024 & 0.9828 & 0.9409  & 0.8298 & 0.9776 & 0.8977 \\
Global cut & 0.9892 & 0.9737 & 0.9814  & 0.9940 & 0.9727 & 0.9832 \\
\underline{Block cut} & 0.9901 & 0.9760 & \textbf{0.9830}  & 0.9948 & 0.9738 & \textbf{0.9842} \\
\hline
Combined best settings & 0.9897 & 0.9827 & \textbf{0.9862}  & 0.9954 & 0.9805 & \textbf{0.9879} \\
Best settings without ethnicity features & 0.9862 & 0.9819 & 0.9841 & 0.9937 & 0.9815 & 0.9876 \\
  \hline
\end{tabular}
\end{table*}
\end{small}

\subsubsection{State-of-the-art.} Most methods proposed in related works have released
neither their software, nor their data, making a fair comparison very difficult.
Yet, we believe solutions reported in the literature can be closely
matched to our generic pipeline, provided the blocking strategy, the linkage
function and the clustering algorithm are properly aligned. In particular,
we consider hereon as the \textit{state-of-the-art} solution the following
combination of components:

@@

As Table~\ref{table:results} shows, switching to either Double metaphone or NYSIIS phonetic-based blocking allows
to improve the overall F-measure score.
In particular, the NYSIIS-based phonetic blocking shows to be the most effective when applied
to the state-of-the-art (with an F-measure of $0.9850$) while also being the most
efficient computationally (with 10857 blocks versus 12978 for the baseline).

@@

Moreover, we observed that the ethnicity features are an improvement. When these features were not
included in the features set, using the best combined settings, the algorithm yields worse result
($0.9841$).

@@

Clustering together all signatures from the same block (\ie, baseline) is the least effective
strategy ($0.9409$), but yields anyhow surprisingly decent accuracy, given the
fact it requires no linkage function and no agglomerative clustering -- only the blocking
function is needed to group signatures. In particular, this result reveals
that author names are not ambiguous in most cases\footnote{This holds for the
data we extracted, but may in the future, with the rise of non-Western
researchers, be an underestimate of the ambiguous cases.} and that only a small
fraction of them requires advanced disambiguation procedures. On the other
hand, both global and block cut thresholding strategies give better results,
with a slight advantage for the block cuts ($0.9814$ versus $0.9830$), as expected.
In case ${\cal S}^\prime_b$ is empty (\ie partial clusters are not known for any of the signatures from the
block), this therefore suggests that either using
a cut-off threshold learned globally from the known data would in general give results only
marginally worse than if the claimed signatures had been
known. Trying out 3 clustering methods, the results show that \textit {average linkage} was the best one.

\longpage

@@

Moreover, the techniques presented in this work can be easily adapted to the broader
problem of named entity disambiguation and might improve accuracy of semantic search
algorithms.
